---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r Created a titanic_data dataframe from csv file}
titanic_data <- read.csv("cleaned_titanic_data.csv")
```
Created a titanic_data dataframe from csv file

```{r Installed and Attached Necessary Packages}
install.packages("tidyverse")
install.packages("plotly")
library(tidyverse)
library(plotly)
```
Installed and Attached Necessary Packages

```{r Bar Plot of Survived}
ggplot(titanic_data, aes(x = as.factor(survived))) +
  geom_bar(fill="purple") +
  xlab("Survived") +
  ylab("Count") +
  ggtitle("Count of Survived Passengers on the Titanic")
```
Bar Plot of Survived

```{r Histogram of Age}
ggplot(titanic_data, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  xlab("Age") +
  ylab("Count") +
  ggtitle("Age Distribution of Passengers")
```
Histogram of Age

```{r Boxplot of Age by Survived}
ggplot(titanic_data, aes(x = as.factor(survived), y = age)) +
  geom_boxplot(fill="orange") +
  xlab("Survived") +
  ylab("Age") +
  ggtitle("Age Distribution by Survival Status")
```
Boxplot of Age by Survived

```{r Violin Plot of Age by Survived}
ggplot(titanic_data, aes(x = as.factor(survived), y = age)) +
  geom_violin(fill="orange") +
  xlab("Survived") +
  ylab("Age") +
  ggtitle("Age Distribution and Density by Survival Status")
```
Violin Plot of Age by Survived

```{r Bar Plot of Passenger Class}
ggplot(titanic_data, aes(x = as.factor(pclass))) +
  geom_bar(fill = "red") +
  xlab("Passenger Class") +
  ylab("Count") +
  ggtitle("Count of Passengers by Class")
```
Bar Plot of Passenger Class

```{r Bar Plot of Embarked}
ggplot(titanic_data, aes(x = embarked)) +
  geom_bar(fill = "purple") +
  xlab("Embarkation Point") +
  ylab("Count") +
  ggtitle("Count of Passengers by Embarkation Point")
```
Bar Plot of Embarked

```{r Scatter Plot of Age vs Fare}
ggplot(titanic_data, aes(x = age, y = fare)) +
  geom_point(color = "orange") +
  xlab("Age") +
  ylab("Fare") +
  ggtitle("Scatter Plot of Age vs. Fare")
```
Scatter Plot of Age vs Fare

```{r Facet Grid of Age vs. Fare by Survived}
ggplot(titanic_data, aes(x = age, y = fare)) +
  geom_point(colour="blue") +
  facet_grid(. ~ survived) +
  xlab("Age") +
  ylab("Fare") +
  ggtitle("Age vs. Fare by Survival Status")
```
Facet Grid of Age vs. Fare by Survived

```{r Facet Grid of Age vs. Fare by pclass}
ggplot(titanic_data, aes(x = age, y = fare)) +
  geom_point(color = "brown") +
  facet_grid(. ~ pclass) +
  xlab("Age") +
  ylab("Fare") +
  ggtitle("Age vs. Fare by Passenger Class")
```
Facet Grid of Age vs. Fare by pclass

```{r Stacked Bar Plot of Survival by Passenger Class}
ggplot(titanic_data, aes(x = pclass, fill = as.factor(survived))) +
  geom_bar(position = "fill") +
  xlab("Passenger Class") +
  ylab("Proportion") +
  labs(fill = "Survived") +
  ggtitle("Survival Proportions by Passenger Class")
```
Stacked Bar Plot of Survival by Passenger Class

```{r Calculate counts and percentages for interactive plot}
titanic_summary <- titanic_data %>%
  group_by(pclass, survived) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100, 
         survival_status = ifelse(survived == 1, "Survived", "Did Not Survive"))
```
Calculate counts and percentages for interactive plot

```{r Convert to data frame}
titanic_summary <- as.data.frame(titanic_summary)
```
Convert to data frame

```{r Create the interactive stacked bar plot}
plot <- plot_ly(titanic_summary, 
                x = ~pclass, 
                y = ~percentage, 
                type = 'bar', 
                color = ~as.factor(survived),
                text = ~paste('Status:', survival_status, '<br>Percentage:', round(percentage, 2), '%'),
                hoverinfo = 'text',
                textposition = 'auto') %>%
  layout(barmode = 'stack',
         xaxis = list(title = 'Passenger Class'),
         yaxis = list(title = 'Percentage'),
         title = 'Survival Proportions by Passenger Class',
         legend = list(title = list(text = 'Survival Status')))

plot
```
Create the interactive stacked bar plot

```{r Create a scatter plot of Age vs. Fare colored by Embarkation}
ggplot_scatter <- ggplot(titanic_data, aes(x = age, y = fare, color = embarked)) +
  geom_point() +
  xlab("Age") +
  ylab("Fare") +
  ggtitle("Scatter Plot of Age vs. Fare by Embarkation") +
  theme_minimal()

ggplot_scatter
```
Create a scatter plot of Age vs. Fare colored by Embarkation and dislplay the plot

```{r Filter the dataframe for passengers who embarked at Southampton}
southampton_data <- titanic_data %>%
  filter(embarked == "S")
```
Filter the dataframe for passengers who embarked at Southampton

```{r Scatter Plot of Age vs Fare for southampton}
ggplot(southampton_data, aes(x = age, y = fare)) +
  geom_point(color = "red") +
  geom_smooth(method = "lm", col = "blue") +
  xlab("Age") +
  ylab("Fare") +
  ggtitle("Scatter Plot of Age vs. Fare for Southampton")
```
Scatter Plot of Age vs Fare

